---
title: 'Math 4317 (Prof. Swiech, S''18): HW #4'
author: "Peter Williams"
date: "4/15/2018"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Section 20  
*A. Prove that if $f$ is defined for $x \geq 0$ by $f(x) = \sqrt{x}$, then $f$ is continuous at every point of its domain.*  

For $f(x) = \sqrt{x},\ \mathcal{D}(f) = \{x \in \mathbb{R}: x \geq 0\}$, let $a \in \mathcal{D}(f)$.

When $a = 0,\ |f(x) - f(a)| = |\sqrt{x} - 0| = \sqrt{x} < \varepsilon$. If we let $\delta(\varepsilon) = \varepsilon^2,$ when $x < \varepsilon^2,\ |f(x)| < \varepsilon$.

When $a \neq 0,\ |f(x) - f(a)| = |\sqrt{x} - \sqrt{a}| = \frac{|\sqrt{x} - \sqrt{a}|}{|\sqrt{x} + \sqrt{a}|} |\sqrt{x} + \sqrt{a}| = \frac{|x - a|}{|\sqrt{x} + \sqrt{a}|} < \frac{|x - a|}{\sqrt{a}} < \varepsilon \implies$ when $|x - a| < \varepsilon \sqrt{a}$, then, $|f(x) - f(a)| < \varepsilon$, thus we can choose $\delta(\varepsilon) = \varepsilon \sqrt{a} \implies$ $f$ is continuous at every point in its domain.

*B. Show that a "polynomial function"; that is, a function $f$ with the form $f(x) = a_n x^n + a_{n-1} x^{n-1} + ... + a_1 x + a_0,\ x \in \mathbb{R}$ is continuous at every point of $\mathbb{R}$.*  

Relying on the properties of algebraic combinations of continuous of functions, we construct $f$ as a combination of continuous functions to show its continuity. Considering the last term of the polynomial function, denoted here, $f_0(x) = a_0,\ f_0(x)$ is a continuous, constant function, since, for any $a \in \mathbb{R}$ we have $|f_0(x) - f_0(a)| = |a_0  - a_0| < \varepsilon = \delta(\varepsilon),\ \varepsilon > 0$. We consider the second to last term of $f$, $a_1x$, as a constant, $a_1$ multiplied by the identity function, denoted, $f_1(x) = x$. Since $f_1(x) = x$, for any real number $a \in \mathbb{R}$, we have $|f_1(x) - f_1(a)| = |x - a| < \varepsilon = \delta(\varepsilon),\ \varepsilon > 0 \implies a_1f_1(x) = a_1x$ is continuous.

Relying on the continuity of $f_1(x) = x$ multiplied by any constant, we can construct higher order terms of $f$ through repeated multiplication of $f_1(x)$, e.g. $a_2 \cdot f_1(x) \cdot f_1(x) = a_2 x^2$ and $a_n\prod_{j=1}^{n}f_1(x) = a_n \cdot f_1(x) \cdot f_1(x) \cdot...\cdot f_1(x) = a_nx^n$, and so on, where each term constructed $a_nx^n$ is continuous on $\mathbb{R}$ since it is constructed via algebraic combinations of continuous functions $\implies f(x) = a_n x^n + a_{n-1} x^{n-1} + ... + a_1 x + a_0,$ is continuous at every point $x \in \mathbb{R}$.  

*E. Let $f$ be the function on $\mathbb{R} \to \mathbb{R}$ defined by $f(x) = x,\ x$ irrational, $f(x) = 1 - x,\ x$ rational. Show that $f$ is continuous at $x = \frac{1}{2}$ and discontinuous elsewhere.*  

Considering the point $a = \frac{1}{2}$, we have $f(a) = \frac{1}{2}$, and $|f(x) - f(a)| = |1 - x - \frac{1}{2}| = |\frac{1}{2} - x| = |x - a| < \varepsilon = \delta(\varepsilon)$. So if $|f(x) - f(a)| < \varepsilon = \delta(\varepsilon) > 0 \implies |x - a| < \delta(\varepsilon)$, and then we have $f$ continuous at the point $a = \frac{1}{2}$. 
For the case $a \neq \frac{1}{2},\ a$ irrational, take a sequence $X = (x_n)$ of rational numbers converging to $a$. Since the sequence $(f(x_n))$ converges to $1 - a$, and we have $f(a) = a$, $f$ is not continuous at irrational points by the Discontinuity Criterion. 
For the case $a \neq \frac{1}{2},\ a$ rational, take a sequence $Y = (Y_n)$ of irrational numbers converging to $a$, the sequence (f(y_n)) converges to $a$, but $f(a) = 1 - a$, which equation is only satisfied when $a = \frac{1}{2}$, thus $f$ is not continuous for rational numbers at any point other than $\frac{1}{2}$.

*F.Let $f$ be continuous on $\mathbb{R} \to \mathbb{R}$. Show that if $f(x) = 0$ for rational $x$, then $f(x) = 0$ for all $x \in \mathbb{R}$.*  

Every real point, $x \in \mathbb{R}$ is the limit of a sequence of rational numbers. If $f$ is continuous $\implies$ for a sequence of rational numbers $X = (x_n) \to x$, we have $(f(x_n)) = 0,$ for all $n \in \mathbb{N}$. Since $f$ is continuous at each rational point $x \in \mathbb{R}$, we can find $|f(x_n) - f(x)| < \varepsilon,\ \varepsilon > 0$, and $|x_n - a| < \delta(\varepsilon) \implies (f(x_n)) \to f(x) = 0, \forall x \in \mathbb{R}$.

*I. Using the results of the preceding exercise, show that the function $g$, defined on $\mathbb{R} \to \mathbb{R}$ by $g(x) = x sin(\frac{1}{x}),\ x \neq 0$, $g(x) = 0,\ x = 0$ is continuous at every point. Sketch a graph of this function.*  

For the case $a = 0$, we have $|g(x) - g(a)| = |x \sin\frac{1}{x} - 0| = |x||\sin\frac{1}{x}| \leq |x| \cdot 1 < \varepsilon,\ \varepsilon > 0$, since $-1 \leq \sin\frac{1}{x} \leq 1$. So when $|g(x) - g(0)| < \varepsilon = \delta(\varepsilon)$, we then have $|x| = |x - 0 | < \delta(\varepsilon) \implies g$ continuous at $0$.

For the case $a \neq 0$, we have $|g(x) - g(a)| = |x \sin \frac{1}{x} - a \sin \frac{1}{a}| = |x \sin \frac{1}{x} - a \sin \frac{1}{a} - a \sin \frac{1}{x} + a \sin \frac{1}{x}| = |(x - a)(\sin \frac{1}{x}) + a(\sin\frac{1}{x} - \sin \frac{1}{a})| \leq |x - a||\sin\frac{1}{x}| + |a||\sin\frac{1}{x} - \sin\frac{1}{a}|$, by Triangle Inequality. Since both $|\sin\frac{1}{x}| \leq 1$ and $|\sin\frac{1}{x} - \sin\frac{1}{a}| \leq 1$, we have $|x - a||\sin\frac{1}{x}| + |a||\sin\frac{1}{x} - \sin\frac{1}{a}| \leq |x - a| \cdot 1 + |a|\cdot 1  = |x - a| + |a| < \varepsilon$. It then follows that if $\delta(\varepsilon) = \varepsilon - |a|$, i.e. $\varepsilon > \delta(\varepsilon) + |a|$, when $|g(x) - g(a)| < \varepsilon$, then $|x - a| < \delta(\varepsilon) \implies$ g continuous at every point in $\mathbb{R}$.  

Sketch of function below:    

```{r gsin,echo=F, fig.height = 4, fig.width= 6}
x <- seq(-.45,.45,by=0.002)
suppressWarnings(y <- ifelse(x == 0, 0, x * sin(1/x)))
plot(x, y, ylab = 'y = f(x) = x * sin (1/x)', bty = 'n', type = 'l')
abline(h =0, lty = 5)
```

*N. Let $g: \mathbb{R} \to \mathbb{R}$ satisfy the relation $g(x + y) = g(x) g(y),\ x,y \in \mathbb{R}$. Show that if $g$ is continuous at $x = 0$, then $g$ is continuous at every point. Also if $g(a) = 0$ for some $a \in \mathbb{R}$, then $g(x) = 0$ for all $x \in \mathbb{R}$.*  

If $g$ is continuous at $x = 0 \implies g(x +y) = g(y) = g(0) \cdot g(y)$. This implies also that $g(0)g(y) = g(y) \implies g(0)g(y) - g(y) = 0 = g(y)(g(0) - 1) = 0 \implies g(0) = 1$, or that $g(0) = 0$.  
If $g(0) = 0 \implies - g(y) = 0 = g(y)$. In this case then $g(y) = 0,\ \forall y \in \mathbb{R} \implies g(x) = 0,\  \forall x \in \mathbb{R}$.  
On the other hand if $g(0) = 1, \implies g(0) \cdot g(y) = g(y)$ continuous for every point $y \in \mathbb{R}$.

#### Section 21
*I. Let $g$ be a linear function from $\mathbb{R}^p \to \mathbb{R}^q$. Show that $g$ is one-one and only if $g(x) = 0$ implies that $x = 0$.*  
Since $g$ is linear $\implies$ for $x,y \in \mathbb{R}^p$, $g(x + y) = g(x) + g(y)$. Then if $g(x) = 0 \implies g(x + y) = 0 + g(y) = g(y) \implies g(x + y) = g(y) \implies g(x + y) = g(0 + y) = g(y)$ which implies $x = 0$. 
If we assume that $g$ is one-one, then for any $g(x) = g(y) \implies x = y$. So in the case $g(x) = 0$, and $g(x + y) = g(x) + g(y) = 0 + g(y)$. Since $g(x) + g(y) = g(y) \implies g(y) - g(x) = g(y) \implies x + y = x - y$, which is satisfied when $x = 0$. 

*J. If $h$ is a one-one linear function from $\mathbb{R}^p \to \mathbb{R}^p$, show that the inverse function $h^{-1}$ is a linear function from $\mathbb{R}^p \to \mathbb{R}^p$.*  

Since $h$ is one-one $\implies$ if $h(x_1) = h(x_2),\ x_1 = x_2,\ x_1, x_2 \in \mathbb{R}^p$. 
Extending the the linear case, we have if $h(ax + by) = h(ax_1 + by_1) = a h(x) + b h(y) = ah(x_1) + bh(y_1)$ then $ax_1 + by_1 = ax + by$. By definition $h^{-1} = \{ax + by: h(ax + by) \in \mathbb{R}^p\} = \{ax : h(ax) \in \mathbb{R}^p\} + \{by : h(by) \in \mathbb{R}^p\}$. This implies $h^{-1}(ax + by) = h^{-1}(h(ax)) + h^{-1}(h(by)) \implies h^{-1}$ is linear, and $h^-1: \mathbb{R}^p \to \mathbb{R}^p$, since  $h^{-1}(h(ax)) + h^{-1}(h(by)) = ax + by \in \mathbb{R}^p$ by construction.  
 
*K. Show that the sum and the composition of two linear functions are linear functions.*  

By definition a function is linear if $f(ax + by) = af(x) + bf(y),\ a,b \in \mathbb{R},\ x,y \in \mathbb{R}^p$. 

For the sum of two linear functions we then have $(f+g)(ax + by) = f(ax + by) + g(ax + by) = af(x) + bf(y) + ag(x) + bf(y) = a(f(x) + g(x)) + b(f(y) + g(y)) = a(f+g)(x) + b(f+g)(y) \implies$ linearity. 
For the composition of two linear functions we have $f \circ g(ax + bx) = f(g(ax + by)) = f(ag(x) + bg(y)) = a f(g(x)) + b f(g(y)) = a(f \circ g)(x) + b(f \circ g)(y) \implies$ composition of two linear functions is linear. 

\newpage

*L. If $f$ is a linear map on $\mathbb{R}^p \to \mathbb{R}^q$, define $||f||_{pq} = \sup \{||f(x)||: x \in \mathbb{R}^p,\ ||x|| \leq 1 \}$. Show that the mapping $f \to ||f||_{pq}$ defines a norm on the vector space $\mathcal{L}(\mathbb{R}^p,\mathbb{R}^q)$ of all linear functions on $\mathbb{R}^p \to \mathbb{R}^q$. Show that $||f(x)|| \leq ||f||_{pq} ||x||$ for all $x \in \mathbb{R}^p$.*  

We have $x = (x_1,x_2,...,x_p) \in \mathbb{R}^p,\ f(x) = y = (y_1,y_2,...,y_q) \in \mathbb{R}^q$, and matrix $A_{q \times p} = (c_{ij}),\ 1 \leq i \leq q,\ 1 \leq j \leq p$, with 
$$y_1 = c_{11}x_1 + x_{12}x_2 + ... + c_{1p}x_p$$  
...  
$$y_q = c_{q1}x_1 + x_{q2}x_2 + ... + c_{qp}x_p$$  
We then have $||f(x)|| = ||(y_1,...,y_q)|| = \sqrt{y_1^2 + ... + y_q^2}$. To show $||f||_{qp} = \sup\{||f(x)||: x \in \mathbb{R}^p,\ ||x|| \leq 1\}$ is a norm in $\mathcal{L}(\mathbb{R}^p,\mathbb{R}^q)$, we have 
(i) $||f||_{pq} \geq 0,\ x \in \mathbb{R}^p$? Since each element in $||f(x)|| = \sqrt{y_1^2 + ... + y_q^2}$, $y_j^2 \geq 0,\ \forall j = 1,...,q \implies \sup \{||f(x)||\} \geq 0 \forall x \in \mathbb{R}^p$ since by definition, $sup\{||f(x)||\} \geq ||f(x)|| \forall x \in \mathbb{R}^p \implies ||f||_{pq} \geq 0$.  
(ii) $||f||_{pq} = 0 \iff f(x) = 0$? Since $||f(x)|| = ||y|| = \sqrt{y_1^2 + ... + y_q^2} = 0 \implies$ each $y_j^2 = 0, \forall j =1,...,q$  
(iii) $\sup ||af(x)|| = |a|\sup ||f(x)|| = |a|||f||_{qp} ,\ a \in \mathbb{R}$? We have $||af(x)|| = ||ay|| = \sqrt{a^2y_1^2 + ... + a^2y_1^2} = \sqrt{a^2}||y|| = |a|||y||$, and $|a| > 0 \implies \sup \{||af(x)||\} = \sup\{|a|||f(x)||\} = |a|sup\{||f(x)||\}$.  
(iv) $\sup\{||f(x + x^{'})||\} \leq \sup ||f(x)|| + \sup||f(x^{'})||,\ x,x^{'} \in \mathbb{R}^p$? Since $f$ is linear $||f(x + x^{'})|| = ||f(x) + f(x^{'})|| \leq ||f(x)|| + ||f(x^{'})||,\ \forall x,x^{'} \in \mathbb{R}^p$ by Triangle Inequality, then $\sup\{||f(x) + f(x^{'})||\} \leq \sup \{||f(x)||\} + \sup\{||f(x^{'})||\}$. 
This implies $||f||_{qp}$ is a norm.  

To show $||f(x)|| \leq ||f||_{pq} ||x||$, we use the earlier notation for a linear map, $f(x) = Ax$, where, $A_{q \times p} = (c_{ij})$. Thus $||f(x)|| = ||Ax|| \leq |A|||x||$ as shown in (21.5). This implies $\sup\{||f(x)||: x \in \mathbb{R}^p, ||x|| \leq 1\} = \sup \{||Ax||\} \leq \sup\{|A|||x||\}$ which is achieved when $x$ is the max value in its domain, i.e. $||x|| = 1$. This implies $\sup\{\||Ax||\}||x|| = \sup\{||f(x)||\} ||x|| = \sup\{||f(x)||\} \cdot 1$. This implies $||f(x)|| \leq \sup\{||f(x)||: x \in \mathbb{R}^p, ||x|| \leq 1\}||x|| \forall x \in \mathbb{R}^p$.

#### Section 22
*B. Let $H: \mathbb{R} \to \mathbb{R}$ be defined by, $h(x) = 1, 0 \leq x \leq 1$. $h(x) = 0, \text{otherwise}$. Exhibit an open set $G$ such that $h^{-1}(G)$ is not open in $\mathbb{R}$, and a closed set $F$, such that $h^{-1}(F)$ is not closed in $\mathbb{R}$.*  

If we take $G = (0,2)$, and open set, $h^{-1}(G) = \{x \in \mathcal{D}(f): h(x) \in G\} = [0,1]$, a closed set. If we take $F = [-2,0]$, a closed set, the inverse image, $h^{-1}(F) = \{x \in \mathcal{D}(f): h(x) \in F\}$  is the union of two open sets $(-\infty,0) \cup (1,+\infty)$ which is open. 

*C. If $f$ is bounded and continuous on $\mathbb{R}^p \to \mathbb{R}$ and if $f(x_0) > 0$, show that $f$ is strictly positive on some neighborhood of $x_0$. Does the same conclusion hold if $f$ is merely continuous at $x_0$?*  

$f$ is bounded and continuous which implies $0 < f(x_0) < M,$ for some $M > 0$. Since $f$ is continuous, for each point $a \in \mathcal{D}(f)$, there is a neighborhood $V$ of $f(a)$ and a neigborhood $U(a) \cap D$ such that if $f(a) \in V \implies a \in U(a)$. Since $f(a) > 0 \implies$ we can take a neighborhood $V$ of $f(a)$ that is also strictly positive, i.e. $V = \{y \in \mathbb{R}: 0 < y < M\}$. If $f$ is not bounded the same argument can be made with $V = \{y \in \mathbb{R}: y > 0\}$. 

*F. A subset $D \subseteq \mathbb{R}^p$ is disconnected if and only if there exists a continuous function $f: D \to \mathbb{R}$ such that $f(D) = \{0,1\}$.*  

$\rightarrow D$ disconnected implies there exists two open sets $B,C$ such that $B \cap D$ and $\ C \cap D$ are disjoint and $(B \cap D) \cup (\ C \cap D) = D$. We can then construct a function $f$ on $D$, $f(x) = 1,\ x \in (B \cap D)$, $f(x) = 0,\ x \in (C \cap D)$.  
$\leftarrow$ Let $f: D \to \mathbb{R}$ be such that $f(D) = \{0,1\} \implies$ the inverse image $f^{-1}(\{0,1\}) = \{ x \in D \subseteq: f(x) \in \{0,1\}\}$ could consist of two disjoint open sets such for $f$ on $D$, $f(x) = 1,\ x \in (B \cap D)$, $f(x) = 0,\ x \in (C \cap D)$, where $D = (B \cap D) \cup (C \cap D) \subseteq \mathcal{D}(f) \implies$ there exists a continuous function $f: D \to \mathbb{R}$ such that $f(D) = \{0,1\}$.

*H. Let $f,g_1,g_2$ be related by the formulas in the preceding exercise. Show that from the continuity of $g_1$ and $g_2$ at $t = 0$ one cannot prove the continuity of $f$ at $(0,0)$.*  

Considering $g_1,g_2$ which are valid are restrictions of the domain of $f$, given $x = (x_1,x_2) \in \mathbb{R}^2$, we can construct $f(x) = 0,\ x_1 \cdot x_2 = 0,\ f(x) = 1,\ x_1 \cdot x_2 \neq 1$. With this $f$ we have $\lim_{x \to (0,0)} \neq 0,$ and $f((0,0)) = 0 \implies$ discontinuity for $f$ at $(0,0)$. Therefore continuity for $g_1,g_2$ on restrictions of $\mathcal{D}(f)$ does not imply continuity of $f$. 

*K. Give an example of a bounded and continuous function $g$ on $\mathbb{R} \to \mathbb{R}$ which does not take on either of the numbers $\sup\{g(x): x \in \mathbb{R}\}$ or $\inf\{g(x): x \in \mathbb{R}\}$*  

If we take $f: \mathbb{R} \to \mathbb{R}$, $f(x) = x,\ x \in (0,1) \subseteq \mathbb{R}$, the function is bounded above by $1$, below by $0$, and continuous on $(0,1)$, but $f(x) \neq 1 = \sup\{f(x): x \in (0,1)\}$, and $f(x) \neq 0 = \inf \{f(x): x \in (0,1)\}$ for any x in interval $(0,1)$.

*O. Let $f$ be a continuous function on $\mathbb{R} \to \mathbb{R}$ which is strictly increasing (in the sense that if $x^{'} < x^{''}$ then $f(x^{'}) < f(x^{''}))$. Prove that $f$ is injective and that its inverse function is continuous and strictly increasing.*  

For points $x,a,b \in \mathcal{D}(f)$, by $f$ be strictly increasing, we have $a > b \implies f(a) > f(b)$, $a = b \implies f(a) = f(b)$ and $a < b \implies f(a) < f(b)$. If we take point $x$ to be $a < x < b$, we can define two neighborhoods $(a, b) \subseteq \mathcal{D}(f)$, and $(f(a),f(b)) \subseteq \mathcal{R}(f)$, such that $x \in (a, b)$, and $f(x) \in (f(a),f(b))$.  This implies $f^{-1}$ in continuous, and since $f^{-1}(f(a)) = a > f^{-1}(f(b)) = b$ if $f(a) > f(b)$, implies $f^{-1}$ is strictly increasing. Also since, $f(a) = f(b) \implies a = b$, $f$ is injective.  

#### Section 23  
*A. Examine each of the functions in Example 20.5 and either show that the function is uniformly continuous on its domain or not.*  
(a) The constant function, $\mathcal{D}(f) \subseteq \mathbb{R}, f(x) = c\ ,\  \forall x \in \mathcal{D}(f)$, where $c$ is a real number.   
Let $\varepsilon > 0$, we have $|f(x) - f(y)| = |0 - 0| = 0 < \varepsilon,\ \forall x,y \in \mathcal{D}(f)$. Regardless of the choice of $\delta(\varepsilon)$, we have $|f(x) - f(y)| < \varepsilon$ whenever $|x - y| < \delta(\varepsilon) \implies$ uniform continuity.  

(b) The identity function $f(x) = x,\ x \in \mathbb{R}$.  
For all $x,y \in \mathbb{R}$, we have $|f(x) - f(y)| < \varepsilon,\ \varepsilon > 0.$ Choose $\delta(\varepsilon) = \varepsilon$. Then whenever $|f(x) - |f(y)| = |x-y|< \varepsilon = \delta(\varepsilon)$ we have $|x- y| < \delta(\varepsilon) \implies$ uniform continuity.    

(c) $f(x) = x^2,\ x \in \mathbb{R}$  
If we take $\varepsilon = 1,$, and consider point positive real points $x, y = x + \frac{1}{2}$, then for $|f(x) - f(y)| = |x^2 - y^2| =|x^2 - (x + \frac{1}{2})^2| = |x^2 - x^2 - x - \frac{1}{4}| = |(-1)(x + \frac{1}{4})| = x + \frac{1}{4} < 1 = \varepsilon$, which is a contradiction, for example, for all $x > 1 \implies f(x)$ not uniformly continuous. 

(e) $f(x) = \frac{1}{x}, x \in \{x \in \mathbb{R}: x \neq 0\}$  
If we take $\varepsilon = 1$, consider points $x,y = \frac{x}{2} \in (0,1) \subseteq \mathbb{R}$ we have $|f(x) - f(y)| = |\frac{1}{x} - \frac{1}{\frac{x}{2}}| = |\frac{-1}{x}| = \frac{1}{x}$. Since both $0< x,y = \frac{x}{2} < 1 \implies \frac{1}{x} > 1$ for all $x,y \in (0,1)$ which implies $f(x) = \frac{1}{x}$ is not uniformly continuous on its domain. 

(f) $f(x) = 0,\ x \geq 0,\ f(x) = 1,\ x > 1$  
Since $f$ is not continuous at point $a=0$, and $a$ is in the domain of $f$, $f$ is not uniformly continuous, since we can find sequence $(f(x_n)) = (f(1/n)) = (1)$ which does not converge to $f(0)$.

(g) $f(x) = 1,\ x\ \text{rational},\ f(x) = 0,\ x\ \text{irrational}$  
$f$ is discontinous at every point in its domain, therefore $f$ cannot be uniformly continuous. 

(h) $\mathcal{D}(f) = \{x \in \mathbb{R}: x > 0\},\ f(x) = 0,\ x\ \text{irrational},\ x > 0$. For rational numbers of the form $\frac{m}{n}$, with $m,n \in \mathbb{N}$ that have no common factor but $1$, $f(\frac{m}{n}) = \frac{1}{n}$.  
Since $f$ is continuous at precisely irrational points, and not all points in its domain, $f$ is not uniformly continuous.  

(i) $\mathcal{D}(f) = \mathbb{R}^2,\ f: \mathbb{R}^2 \to \mathbb{R}^2,\ f(x,y) = (2x + y, x - 3y)$.  For $(x,y),(a,b) \in \mathcal{D}(f)$ we have $||f(x,y) - f(a,b)|| = ||(2x + y -2a -b,x-3y-a+3b)|| \leq \sqrt{2}\sup \{||2x + y -2a -b  ||,||x - 3y - a + 3b||\} \leq \sqrt{2} \cdot 4 ||(x,y) - (a,b)|| \leq \varepsilon$, since $|x - a| \leq \sqrt{(x - a)^2 + (y-b)^2} = ||(x,y) - (a,b)|| \implies ||2x + y - 2a -b|| = ||2(x -a) + (y-b)|| \leq 3 ||(x,y) - (a,b)||$, and since $||x - 3y - a + 3b|| = ||(x -a) + 3(b-y)|| \leq 4 ||(x,y) - (a,b)||$.
Putting this together, we have for $\varepsilon >0$, whenever $||f(x,y) - f(a,b)|| < \varepsilon$ we have $||(x,y) - (a,b)|| \leq \frac{\varepsilon}{4 \sqrt{2}}$ which implies uniform continuity. 

(j) $\mathcal{D}(f) = \mathbb{R}^2,\ f: \mathbb{R}^2 \to \mathbb{R}^2,\ f(x,y) = (x^2 + y^2, 2xy)$.  

*C. If $B$ is bounded in $\mathbb{R}^p$ and $f: B \to \mathbb{R}^p$ is uniformly continuous, show that $f$ is bounded on $B$. Show that this conclusion fails if $B$ is not bounded in $\mathbb{R}^p$.*  
If we take two sequences in $B$, $x_n, y_n,\ n \in \mathbb{N}$ by uniform continuity of $f$, whenever $||x_n - y_n|| \leq \frac{1}{n}$ we have $||f(x_n) - f(y_n)|| < \varepsilon$ for some $\varepsilon > 0$. If we consider the point $x_0 \in B$, for which $f(x_0) = M = \sup \{||f(x)||: x \in B\}$. By Bolzano-Weierstrass, we can find a subsequence of $(x_n)$, $(x_{n1},...,x_{nk})$that converges to $x_0 \implies$ whenever $||x_0 - y_n|| \leq \frac{1}{n}, n \in \mathbb{N}$, we have $||f(x_0) - f(y_n)|| \leq \varepsilon \implies f$ is bounded on $B$.  

*D. Show that $f(x) = \frac{1}{1 + x^2}$ for $x \in \mathbb{R}$ is uniformly continuous.*   
Take $\varepsilon > 0$, for $x,y \in \mathbb{R}$, we have $|f(x) - f(y)| = |\frac{1}{1 + x^2} - \frac{1}{1 + y^2}| = |\frac{(1 + y^2)}{(1 + x^2)(1 + y^2)} - \frac{(1+ x^2)}{(1 + y^2)(1 + x^2)}| = |\frac{y^2 - x^2}{(1 + y^2)(1 + x^2)}| = |x+y||x-y||\frac{1}{(1+x^2)(1 + y^2)}| \leq (|\frac{x}{(1+x^2)(1+y^2)}| + |\frac{y}{(1+x^2)(1+y^2)}|) |x-y| \leq (|\frac{y}{(1+y^2)}| + |\frac{x}{(1+x^2)}|   )|x - y|$. Since $\forall x \in \mathbb{R}$, we have $|\frac{x}{1 + x^2}| <1$, we have $(|\frac{y}{(1+y^2)}| + |\frac{x}{(1+x^2)}|   )|x - y| < 2|x-y| = 2\delta(\varepsilon) \implies$ if we choose $\delta(\varepsilon) = \frac{\varepsilon}{2}$ whenever $|x - y| < \frac{\varepsilon}{2}$ we have $|f(x) - f(y)| < \varepsilon$, for all $x,y \in \mathbb{R}$

*F. Show that $f(x) = \frac{1}{x^2},\ \mathcal{D}(f) = \{x \in \mathbb{R}: x > 0\}$ is not uniformly continuous on its domain.*  
If we take $\delta(\varepsilon) = \varepsilon/2,\ \text{and}\ \varepsilon = 1$, and consider points in a subset of $\mathcal{D}(f)$, namely $x,y \in (0,1)$,and then take $y = \frac{x}{2} \in (0,1)$, we have we have $|x - y| = |x/2| < \varepsilon/2 = 1/2 \implies |f(x) - f(y)| = |\frac{1}{x^2} - \frac{4}{x^2}| = \frac{3}{x^2} < 1$. But, for all $x, y \in (0,1),\ \frac{3}{x^2} > 1 \implies f(x)$ is not uniformly continuous on its domain. 

*G. A function $g:\mathbb{R} \to \mathbb{R}^p$ is periodic if there exists a number $p>0$ such that $g(x + p) = g(x)$ for all $x \in \mathbb{R}$. Show that a continuous periodic function is bounded and uniformly continuous on $\mathbb{R}$.*   
We assume $g$ is continous, it implies for $x \in \mathbb{R}$, if we consider points $x,y$ over the domain/interval $[x_0,x_0+p]$ whenever $|x - y| \leq |x_0 - x_0 - p| = |-p| = p = \delta > 0$ we have $|g(x) - g(y)| < \varepsilon$. Considering points $x + np, y + np,\ n \in \mathbb{N}$, we have $|(x + np)  - (y + np)| = |x - y| < \delta$ implying that $|g(x + np) - g(y + np)| = |g(x) - g(y)| < \varepsilon \implies g$ is bounded and uniformly continous on $\mathbb{R}$. 

*H. Let $f$ be defined on $D \subseteq \mathbb{R}^p$ to $\mathbb{R}^q$, and suppose that $f$ is uniformly continuous on $D$. If $(x_n)$ is a Cauchy sequence in $D$, show that (f(x_n)) is a Cauchy sequence in $\mathbb{R}^q$.*  
A sequence is Cauchy if for some $\delta >0\ \exists M \in \mathbb{N}$ such that for all $m,n \geq M$ then we $||x_m - x_n|| < \delta$. Since $f$ is uniform continuous, for $||f(x_m) - f(x_n)||$, for $x_m, x_n \in D$ for all $m,n \in \mathbb{N}$, whenever $||x_m - x_n|| < \delta$, we have $||f(x_m) - f(x_n)|| \leq \varepsilon$ for some $\varepsilon > 0 \implies$ there exists some $M \in \mathbb{N}$ such that for all $m,n \geq M$, $||f(x_m) - f(x_n)|| \leq \varepsilon \implies (f(x_n))$ is Cauchy.

#### Section 24  
*B. Give an example of a sequence of everywhere discontinuous functions which converges uniformly to a continuous function.*    
*D. Let $(f_n)$ be a sequence of continuous functions on $D \subseteq \mathbb{R}^p$ to $\mathbb{R}^q$ such that $(f_n)$ converges uniformly to $f$ on $D$, and let $(x_n)$ be a sequence of elements in $D$ which converges to $x \in D$. Does it follow that $(f_n(x_n))$ converges to $f(x)$?*  
Since each $f_n, n \in N$ is continuous, $f$ is continuous. Then whenever, $||x_n - x|| < \delta$, for some $n \geq K \in \mathbb{N}$ we can take $||f(x_n) - f(x)|| < \frac{\varepsilon}{2}$, for some $\varepsilon > 0$. Considering the sequence $f_n$, we have $||f(x_n) - f(x)|| = ||f(x_n) - f_n(x_n) + f_n(x_n) - f(x)|| \leq ||f(x_n) - f_n(x_n)|| + ||f_n(x_n) - f(x)|| = ||f_n(x_n) - f(x_n)|| + ||f_n(x_n) - f(x)||$. If we take $n \geq M \in \mathbb{N}$, $||f_n(x_n) - f(x_n)|| \leq \frac{\varepsilon}{2}$, by the uniform continuity of $f_n$. This implies that for $n \geq \sup \{K,M\}$ we have $||f_n(x_n) - f(x_n)|| + ||f_n(x_n) - f(x)|| \leq \frac{\varepsilon}{2} + \frac{\varepsilon}{2} = \varepsilon \implies ||f_n(x_n) - f(x)|| \leq \varepsilon \implies f_n(x_n) \to f(x)$.
*E. Consider the sequences $(f_n)$ defined on $D = \{x \in \mathbb{R}: x \geq 0\}$ to $\mathbb{R}$ by the following formulas. Discuss the convergence and uniform convergence of these sequences and the continuity of the limit functions. In case of non-uniform convergence consider appropriate intervals in $D$.*  
(b) $\frac{x^n}{1 + x^n}$,  
For $0 \leq x < 1$, we have $f_n(x) = \frac{x^n}{1 + x^n} \to_{n \to \infty} 0$ since $x^n \to 0$ for $0\leq x<1$. For $x = 1$, $f_n(x) = \frac{x^n}{1 + x^n} = \frac{1}{2},\ \forall n \in \mathbb{N}$. For $x > 1$, $f_n(x) = \frac{x^n}{1 + x^n} \to_{n \to \infty} 1$ which implies $(f_n)$ is pointwise convergent. 
To examine uniform convergence, we have limit function $f(x) = 0,\ 0 \leq x < 1$, $f(x) = \frac{1}{2},\ x = 1$, and then $f(x) = 1, x > 1 \implies$ uniform converges on closed intervals falling within the interval $x > 1$, or within the interval $0 \leq x < 1$, but not for closed intervals containing the point $1$, since the limit function, for example, for $x$ approaching $1$ from below, $\lim f_n(x) = 0$, but $f_n(1) = 1/2, \forall n \in \mathbb{N}$.  We then do not have uniform convergence over the entire domain, given discontinuous limit functions. 

(c) $\frac{x^n}{n + x^n}$,  
For $0 \leq x <1$, we have $f_n(x)  = \frac{x^n}{n + x^n} \to \frac{0}{n + 0} \to 0$. For $x = 1$, we have $f_n(1) = \frac{1}{n + 1} \to 0$. And for $x > 1$, we have $f_n(x) = \frac{x^n}{n + x^n} = \frac{\frac{x^n}{n}}{1 + \frac{x^n}{n}} \to 1$, which implies pointwise convergence over $x \geq 0$. To examine uniform convergence we have limit function $f(x) = 0$ for $0 \leq x \leq 1$, and then $f(x) = 1$ for $x > 1$. For $x \in [0,1]$, we have $||f_n - f||_D = \sup \{||x^n/(n + x^n)||: x \in [0,1]\} = \frac{1}{n+1} \to_{n \to \infty} 0 \implies$ uniform continuity on interval $[0,1]$. For $x > 1$, we have $||f_n - f||_D = \sup \{||\frac{x^n}{n + x^n} - 1||: x > 1\}$, and $||\frac{x^n}{n + x^n} - 1|| = ||\frac{x^n}{n + x^n} - \frac{n + x^n}{n + x^n}|| = ||\frac{-n}{n + x^n}|| = \frac{n}{n + x^n} = \frac{1/n}{(1/n) + (x^n/n)} \to 0$, since $x > 1$. This implies uniform convergence on the interval $x \in [a,\infty)$, such that $a > 1$.  

(d) $\frac{x^{2n}}{1 + x^n}$,  
For $0 \leq x <1$, we have $f_n(x)  = \frac{x^{2n}}{1 + x^n} \to \frac{0}{1 + 0} \to 0 = f(x)$. For $x = 1$, we have $f_n(1) = \frac{1}{1 + 1} \to \frac{1}{2} = f(1)$. And for $x > 1$, we have $f_n(x) = \frac{x^{2n}}{1 + x^n} = \frac{x^{2n}/n}{1/n + x^n/n} \to \frac{x^{2n}/n}{x^n/n} \to x^n \to +\infty$, which implies pointwise convergence over the first two intervals, $0 \leq x < 1$, and $x = 1$.  
To examine uniform convergence, we have limit function $f(x) = 0,\ 0 \leq x < 1$, $f(x) = \frac{1}{2},\ x = 1$, and then divergence for $x > 1$. This implies uniform converges on closed intervals falling within the interval $0 \leq x < 1$, but including the point $1$, since the limit function, for example, for $x$ approaching $1$ from below, $\lim f_n(x) = 0$, but $f_n(1) = 1/2, \forall n \in \mathbb{N}$. For $x >1$ we have a divergent sequence of functions. We then do not have uniform convergence given discontinuous limit functions. 

(e) $\frac{x^n}{1 + x^{2n}}$  
For $0 \leq x <1$, we have $f_n(x)  = \frac{x^{n}}{1 + x^{2n}} \to \frac{0}{1 + 0} \to 0 = f(x)$. For $x = 1$, we have $f_n(1) = \frac{1}{1 + 1} \to \frac{1}{2} = f(1)$. And for $x > 1$, we have $f_n(x) = \frac{x^n}{1 + x^{2n}} = \frac{x^{n}/n}{1/n + x^{2n}/n} \to \frac{x^n/n}{x^{2n}/n} = \frac{1}{x^n} \to 0$, which implies pointwise convergence over the first two intervals, $0 \leq x < 1$, and $x = 1$.  
To examine uniform convergence, we have limit function $f(x) = 0,\ 0 \leq x < 1$, $f(x) = \frac{1}{2},\ x = 1$, and then $f(x) = 0$ for $x > 1$. For $0 \leq x < 1$, we have $||f_n - f||_D = \sup \{\frac{x^n}{1 + x^{2n}}: 0 \leq x < 1\} = 0 \implies$ uniform convergence on closed intervals contained in clopen interval $[0,1)$. For $x > 1$, we have $||f_n - f||_D = \sup \{\frac{x^n}{1 + x^{2n}}: x > 1\} = 0 \implies$ uniform convergence on closed intervals contained interval $[a,\infty)$, such that $a > 1$. For $x = 1$ we have $f_n(1) = 1/2$, and $f(1) = 1/2$, and thus have discontinuous limit functions. 

*J. Prove the following theorem of G. Polya. If for each $n \in \mathbb{N}$ the function $f_n$ on $I \to \mathbb{R}$ is monotone increasing and if $f(x) = \lim(f_n(x))$ is continuous on $I$, then the convergence is uniform on $I$. (Observe that it is not assumed that $f_n$ is continuous.)*  
$f$ monotone increasing is given. Since $f$ is uniformly continuous, if $\varepsilon > 0$, let $0 = x_0 < x_1 < ... < x_h = 1$ be such that $f(x_j) - f(x_{j-1}) < \varepsilon$ and let $n_j$ be such that if $n \geq n_j$; then $|f(x_j) - f_n(x_j)| < \varepsilon$. If $n \geq \sup\{n_0, n_1, ... ,n_h\}$, show that $|f(x_0) - f_n(x)| < 3 \varepsilon,\  \forall x \in \mathbb{I}$

*N. If $f_3(x) = x^3$ for $x \in \mathcal{I}$, calculate the $n^{th}$ Bernstein polynomial for $f_3$. Show directly that this sequence of polynomials converges uniformly to $f_3$ on $\mathbb{I}$.*  
For $f_3: [0,1] \to \mathbb{R}$, to calculate $B_n(x;f_3)$, for $n = n - 3, k = j$, we have $1 = \sum_{j=0}^{n-3} {n - 3 \choose j} x^j (1 - x)^{n - (j + 3)}$ This $x^3 = \sum_{j=0}^{n-3} {n-3 \choose j} x^{j+3} (1 - x)^{n - (j + 3)} = \sum_{j=0}^{n-3} \frac{(j + 3)(j + 2)(j+1)}{n (n-1)(n-2)} {n \choose j + 3} x^{j+3} (1 - x)^{n - (j + 3)}$. If we let $k = j + 3$, we then have $x^3 = \sum_{k=0}^{n} \frac{(k)(k-1)(k-2)}{n (n-1)(n-2)} {n \choose j + 3} x^{k} (1 - x)^{n - k}$, multiplying through by $\frac{1}{n^3}$, we have $\frac{1}{n^3}n(n-1)(n-2) x^3 = \sum_{k=0}^{n} \frac{k^3 - 3k^2 + 2k}{n^3} {n \choose j + 3} x^{k} (1 - x)^{n - k} = \sum_{k=0}^{n} \frac{k^3}{n^3} {n \choose j + 3} x^{k} (1 - x)^{n - k} - \frac{3}{n}[(1 - \frac{1}{n})x^2 + \frac{1}{n} x] + \frac{2}{n^2}x$, since we have from (24.6), $x = \sum_{j=0}^{n}\frac{k}{n}{n \choose k}x^{k} (1 - x)^{n - k}$, and $(1 - \frac{1}{n})x^2 + \frac{1}{n} x = \sum_{j=0}^{n}\frac{k^2}{n^2}{n \choose k}x^{k} (1 - x)^{n - k}$. We then have $\sum_{k=0}^{n} \frac{k^3}{n^3} {n \choose j + 3} x^{k} (1 - x)^{n - k} = B_n(x;f_3) = \frac{n(n-1)(n-2)x^3}{n^3} + \frac{3x^2(n-1)}{n^2} + \frac{x^2}{n}$. 
By Bernstein approximation theorem, for $f_3(x) = x^3$, we have $\sup\{|f_3 - B_n(x; f_3)|: x \in [0,1]\} \leq 1 * (1 - \frac{n(n-1)(n-3)}{n^3}) + \frac{3}{n} + \frac{1}{n} \to_{n \to \infty} 0$, which implies unfirom continuity on $[0,1]$.

*S. Show that the Weierstrass Approximation Theorem fails for bounded open intervals.*  

Take $(a,b)$ to be an open bounded interval, with $b > a$. The function $f(x) = \frac{1}{b-x},\ x \in (a,b)$, we have $||f(x) - P_n(x)||_D = \sup\{||\frac{1}{b-x} - P_n(x)||: x  \in (a,b)\} = \infty$, since $P_n(x)$ must be bounded on $(a,b)$, and $f(x)$ is unbounded as $x \to b$.

#### Section 26  
*N. If $K \subseteq \mathbb{R}^p$ is compact and $(f_n)$ is a sequence of continuous functions on $K$ to $\mathbb{R}^q$ which is uniformly convergent on $K$, show that the family $\{f_n\}$ is uniformly equicontinuous on $K$ in the sense of Definition 26.6.*  

*O. Let $\mathcal{F}$ be a bounded and uniformly equicontinuous collection of functions on $D \subseteq \mathbb{R}^p$ to $\mathbb{R}$ and let $f^{\*}$ be defined on $D \to \mathbb{R}$ by $f^{\*} = \sup \{f(x): f \in \mathcal{F}\}$. Show that $f^{\*}$ is continuous on $D \to \mathbb{R}$.*  

*Q. Consider the following sequences of functions which show that the Arzela-Ascoli Theorem 26.7 may fail if the various hypotheses are dropped.*  
(a) $f_n(x)  = x + n,\ x \in [0,1]$;  Domain compact, sequence uniformly equicontinuous but not bounded
(b) $f_n(x)  = x^n,\ x \in [0,1]$;  Domain compact, sequence bounded by not uniformly equicontinuous.
(c) $f_n(x) = \frac{1}{1 + (x -n )^2},\ x \in [0,+\infty)$.  Domain not compact, sequence bounded, and uniformly continuous

