---
output:
  pdf_document: default
  html_document: default
  word_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
\section*{Predictive Models for Audiences\\}

Television programmers who are tasked with scheduling decisions are specifically interested the impact of manipulating the ordering of programs within a 24 hour cable network schedule. This consists of determining the sequence of program airing on a television network. Dynamically generating accurate predictions of the expected audience associated with possible permutations of a TV schedule is critical to the decision making process of scheduling programs. Key business requirements for predicting audiences in the programming scheduling context include: flexibility in generating predictions for program in new time slots, the ability to rapidly retrieve audience predictions given schedule input changes, and the capability of interpreting the results of scheduling scenarios. 

Multilevel regression models provide flexibility to the data analyst program level audience variation can be explicity accounted for when predicting audience levels when the ordering or sequence of programs on the TV schedule change. Here we briefly present exemplary multilevel regression models employed to support the task of predicting audiences. The data employed consists of observations of average audience impressions, which is a projection of the average number of household viewing a specific TV network in a specific time period. Typically the data may be organized by hour, or half hour, so that for each date, and TV network we observe 24 or 48 observations of average audience. Each of these observations are associated with a specific TV program, that was airing for the majority or entirety of the time period (hour, or half hour). For an exemplary model, we then have \text{for observations,} i = 1,2,3,...,N\, \text{programs, } j = 1,2,3,...,J, we have: 

\begin{equation}
  y_{i} \sim N(\alpha_{j[i]} + \beta_{1j[i]}x_{1i} + \beta_{2}x_{2i} + \varepsilon_i),\ \varepsilon_i \sim N(0, \sigma_y^2)\ \text{where,}  
\end{equation}

\begin{equation}
  \begin{pmatrix}
  \alpha_j \\
  \beta_j
  \end{pmatrix}  \sim
  N(\begin{pmatrix} 
  \mu_{\alpha} \\ 
  \mu_{\beta_1} 
  \end{pmatrix},
  \begin{pmatrix} 
\sigma_{\alpha}^2 & \rho \sigma_{\alpha} \sigma_{\beta_1} \\ 
\rho \sigma_{\alpha} \sigma_{\beta_1} &  \sigma_{\beta_1}^2 \end{pmatrix})
\end{equation}

Here the index notation j[i] associates an observation with its corresponding specific TV program, for example, 1[2] indicates that the second observation, is associated with the program possessing the first index value. We assume, as is always the case practically, N > J, i.e. there are more observations than program, or in other words, we have repeated observations associated with the same program.


We utilize xx different variables, measured at xx different levels, noted below by subscript pas fixed-effects predictors including, day of week (7  1 indicator variables), an indicator variable ofwhether the program is a repeat, and indicator of whether the program on the schedule is leadingout of a program of the same genre type (e.g.  situation comedy, movie, variety show etc.)(1), anindicator program is leading out of a movie (1), quarter of year (4  1 indicator variables), and stackorder  (a  numeric  variable  which  counts  the  number  of  sequential  hours  the  program  airs).   Therandom component of the model includes intercepts for programs, and hour of the day, the two key interacting factors in the organization of a program schedule, and the main interaction of interestfor those scheduling programs

When applied, this model capability greatly simplifies the decision making process for those tasked with comparing scheduling possibilities. As demonstrated through cross-validation, pooling together audience data of existing programs in current schedule time slots with programs currently airing elsewhere on the schedule is key to this flexibility. Important predictors of audience levels in our analysis of program scheduling include the cable network, the programs that air on the network and the specific time slot, or daypart (at a higher level) they air in. Other variables including the chosen sequencing of programming, genre, along with seasonal and cyclical variable also are beneficial to prediction.Multilevel models provide a methodology to generate granular, telecast level predictions while also accounting for program level variation within a network. For the purpose of program scheduling, the key grouping variables are programs, and day-parts, or time periods during the day when viewing levels are arbitrarily assumed to be homogeneous (e.g. Overnight, Morning, Afternoon, Prime). 

Television programmers are specifically interested in obtaining accurate predictions of audience rating when manipulating the ordering of programs within a 24 hour cable network schedule. Multilevel models are flexible in generating predicting new programs in new time periods and thus help the TV programmer weigh the potential effects on network ratings of changing the ordering of programs.
We utilize eight different variables, measured at 13 different levels, noted below by subscript pas fixed-effects predictors including, day of week (7  1 indicator variables), an indicator variable ofwhether the program is a repeat, and indicator of whether the program on the schedule is leadingout of a program of the same genre type (e.g.  situation comedy, movie, variety show etc.)(1), anindicator program is leading out of a movie (1), quarter of year (4  1 indicator variables), and stackorder  (a  numeric  variable  which  counts  the  number  of  sequential  hours  the  program  airs).   Therandom component of the model includes intercepts for programs, and hour of the day, the two keyinteracting factors in the organization of a program schedule, and the main interaction of interestfor those scheduling programs

The long term model is a multilevel regression is implemented where the observations are represented at a date, half-hour level, indexed $i=1,2,...,n$. Average audience at the observation level, are the the response of interest in forecasting. In practice, the response is represented transformed $y_i = log(AA_i + 1)$, where $AA_i$ representes the average audience or impressions at the observation level, This choice of transform is informed by the occurence of natural zeroes, that is, observations with no impressions delivered.\\ 
An exemplary model for the transformed average audience is written below. The model is a basic multilevel regression with predictors that vary at the television program-level, indexed $j = 1,...,J$, and predictors that are common across all all observations $i=1,..,n$. We can extend previous notation to matrix representation below

\begin{equation}
\begin{split}
  y_{i} = X_{i}^{0}\beta^{0} + X_{i}B_{j[i]} + \epsilon_{i}, \: \text{for}\: i = 1,...,n \:\: \text{ with, }\\
  y_{i} \sim N(X_{i}^{0}\beta^{0} + X_{i}B_{j[i]}, \sigma_{y}^{2}),\: \text{  and,     }\\ 
  B_{j} \sim N(U_{j}G, \Sigma_{B}), \: \text{,     for}\: j = 1,...,J
\end{split}
\end{equation}

where:
\begin{itemize}
  \item $X$ is a $J \times 2$ matrix of program-level predictors and $B$ a $2 \times J$ matrix of program-level coefficients. This enables a program-level intercept and slope on the PUT Factor, where $j[i]$ represents the program associated with observation $i$ that represents a half hour program airing specific to a date and time.
  
  \item $U$ is a $J \times 2$ matrix of program level predictors and $G$ is a $2 \times 2$ matrix of coefficients at the program level. Their product is a vector of length $2$ that is modeled as a bivariate normal, with covariance matrix $\Sigma_{B}$. 
  
  \item $X^{0}$ is a $n \times L$ matrix of predictors and $\beta^{0}$ a $L \times 1$ vector of regression coefficients - common to all observations. These predictors include binary or dummy variables identifying whether an individual observation represents a repeat or new telecast of an episode of a television program, a variable that measures the historical counts of programs of the same genre, a combination of trignometric functions to capture seasonality in television —(cite Danaher, 2012 and (Gensch and Shaman 1980):. 
  
\end{itemize}
To estimate model parameters, the REML or restricted-maximum likelihood crierion is used see (Bates 2014) for more details. 
% @book{gelman2006data,
%   title={Data analysis using regression and multilevel/hierarchical models},
%   author={Gelman, Andrew and Hill, Jennifer},
%   year={2006},
%   publisher={Cambridge university press}
% }
% @article{bates2014fitting,
%   title={Fitting linear mixed-effects models using lme4},
%   author={Bates, Douglas and M{\"a}chler, Martin and Bolker, Ben and Walker, Steve},
%   journal={arXiv preprint arXiv:1406.5823},
%   year={2014}
% }
\end{document}
```{r, include = F}
#lmer(bc_p2554 ~ (year_seq + scaled_put + stackc_log + repeat_factor|prg_factor) + year_seq + year_seq2 + scaled_put + wknd + dur_hour + stackc_log + repeat_factor, data=mdat,REML=T);
```